Template for indicator documentation

### 2023-03-13 slitasje.Rmd
trying to mask

This is a quite heavy operation and should be done on a server.

```{r, eval=F}
#takes about one minute
mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=F)
```

This is in the wrong CRS.

```{r}
mireET <- sf::st_transform(mireET, 25833)
```

```{r, eval=F}
system.time(
  mireET_warped <- stars::st_warp(mireET, grid)
)
```

user: 597.308

system: 139.164

elapsed: 735.907

```{r}
mireET
```

```{r, eval=F}
write_stars(mireET_warped, paste0(pData, "wetland_ecosystem_EPSG25833_50m.tif"))
```

```{r}
wetlands_trd <- sf::st_crop(mireET, myBB)
tm_shape(wetlands_trd)+
  tm_raster()+
  tm_shape(hw_utm)+
  tm_lines(col="blue")
```

## OLD
Instead I will draw some random point samples.

```{r, eval=F}
wetland_points <- st_sample(mireET, 10000)
```

And intersect this with the extrapolated indicator values.

```{r, eval=F}
wetland_points_match <- st_intersection(wetlands_slitasje_extr, wetland_points)
```

And extract the values

```{r, eval=F}
wetland_points_extracted <- st_intersection(wetlands_slitasje_extr, wetland_points)
```

If we went for option 2, we could instead do something like this, and

create an empty grid over Norway, 50 x 50 m resolution

<!--# This we should perhaps do once, in a separate misc chapter, and make available for all chapters -->

```{r, eval=F}
grd = st_as_stars(st_bbox(wetlands_slitasje_extr), dx = 50, dy = 50, values = NA_real_)
```

and then rasterize the indicator

```{r, eval=F}
wetlands_slitasje_extr_rast <- st_rasterize(wetlands_slitasje_extr, grd)
```

To get a perfect alignment with the ET map, we should change `grd` with
the ET raster map.

Then we can use starts filtering with `[ .. ]` to mask out the areas
that do not belong to the ET.

>>>>

```{r, eval=F}
#takes about one minute without the proxy
mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=T)
mireET
```

This is in the wrong CRS. We want to stay with 25833.

We have a 50x50 m [master grid](#master-grid) with the correct CRS, and I would warp mireET to fit that grid like this:
```{r, eval=F}
grid <- stars::read_stars(paste0(pData, "masterGrid_50m.tiff"))
mireET_warped <- stars::st_warp(mireET, grid)
```

... but I don't want to down sample the mire data, since we will loose a lot of the smaller ecosystem occurrences.

I could make a 10x10 master grid, but I have not found a way to stop this from being just a massive file.

I can transform the ET map like this, and not worry about the rasters being perfectly aligned, but even this operation is too heavy for the R Studio Server:

```{r, eval=F}
mireET_trans <- sf::st_transform(mireET, 25833)
```

So I will transform the extrapolated indicator map ti match the ET instead. Note that the ET map is just for southern Norway up to about the middle of the country.

```{r}
system.time(
  wetlands_slitasje_extr <- st_transform(wetlands_slitasje_extr, st_crs(mireET))
)
```

Then I import the real ET data and mask

```{r, eval=F}
mireET <- stars::read_stars("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif",
                             proxy=F)
```

<!-- PLAN: rasterize wetland_slitasje_extr and c() with mireET, and mask -->
<!-- Or use  -->
<!-- wetlands_slitasje_extr[mireET==0] <- NA -->

```{r}
wetlands_slitasje_extr_stars <- stars::st_as_stars(wetlands_slitasje_extr)
temp <- c(wetlands_slitasje_extr_masked, mireET)
class(wetlands_slitasje_extr)
dim(temp)
```
